Hydrodynamics of a Semipermeable Vesicle Under Flow and Confinement

### 2021-01-10 vesicle_code/docs/permeableVes/main.tex
added tension plots
yny

% In this work we use numerical simulations to examine the hydrodynamics
% of a semipermeable vesicle in two dimensions. 
In microfluidic experiments, it is challenging to keep a freely
suspended red blood cell or vesicle in a steady flow for more than a few
hours. It is also difficult to quantify the vesicle area/volume ratio
accurately.  Using numerical simulations we can quantify the effects of
water permeability on the hydrodynamics of a vesicle under mechanical
stresses. Vesicles (self-enclosed lipid bilayer membranes) have been
used as a model system to study reshaping, remodeling, and scission of
cell membranes due to osmotic stress~\cite{OgleckaEtAl2014_eLife,
CamposSaric2020_bioRxiv, Dimova2020_GVB, BhatiaChrist2020_SoftMatt}.
Under an osmotic shock, a vesicle may undergo oscillatory variation in
its enclosed volume~\cite{ChabanonHoLiedberg2017_BJ}. When trapped in a
microfluidic channel, an osmotic shock induces vesicle shape changes,
tank-treading and tumbling
motions~\cite{BhatiaRobinsonDimova2020_SoftMatt,
BhatiaChrist2020_SoftMatt}, an illustration that the permeable water
flow is inevitably coupled with the mechanical pressure jump across the
membrane~\cite{yao-mor2017}, which could become comparable to osmotic
stress when the membrane deformation is large~\cite{LiMoriSun2015_PRL,
yao-mor2017}. In the absence of an osmotic gradient, we first show that
a freely suspended semipermeable vesicle behaves differently from an
impermeable vesicle over a long time. Next we show that, at normal
physiological conditions, the permeable water flow can be amplified by
extreme confinement and results in observable change in both vesicle
hydrodynamics and reduced area. These results will shed light on the
dynamic aging process of living cells that constantly go through flows
and constrictions over a period of days.

@@

In experiments, vesicles under shear flows are measured for a minute
(chapter 19 in~\cite{Dimova2020_GVB}), a duration over which vesicle can
be assumed impermeable~\cite{AbkarianViallat2005_BJ}. Our simulations of
a semipermeable vesicle in these flows show that water permeability may
lead to different vesicle hydrodynamics over a time scale of 10 to 100
minutes. 
%{\bf YNY: is this correct? The shear flow results seem to suggest that
%the permeability may lead to change in area in less than a second???}
For vesicles in a narrow microfluidic channel, the typical time scale is
milliseconds~\cite{abk-fai-sto2006} \todo[inline]{Howard talked about
this statement not always being true}. Our simulations show that strong
confinement can lead to amplification of water permeability at such
short time scale.
%
%all vesicles have a length
%$L$, but the area is not conserved because of semipermeability. We vary
%the initial vesicle reduced area, the permeability rate, and the flow
%rate. For all simulations, the bending modulus is $10^{-19}J$.

@@

Semipermeability is introduced by modifying the vesicle velocity from
the standard no-slip condition $\dot{\xx} = \uu(\xx)$. Instead, we
assume that the flux through the vesicle is proportional to the
mechanical force~\cite{yao-mor2017}. That is, 
\begin{align}
  \label{eqn:vesicleFlux}
  \uu - \dot{\xx} = - k_w (\ff_\mathrm{mem} \cdot \nn) \nn, \qquad
  \xx \in \gamma,
\end{align}
where $k_w$ is the water permeability.

@@

We next compute the transient area of the vesicle. Since the area of the
vesicle is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds,
  \label{eqn:area}
\end{align}
its derivative includes terms due to the time derivative of $\xx \cdot
\nn$ and a term due to interface stretching~\cite{lai-tse-hua2008}, but
this latter term term is zero because of the local inextensibility
condition. Therefore,
\begin{align}
  \dot{A}(t) =
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
Applying integration by parts, the two integrals agree, and we have
%\begin{align}
%  \int_\gamma (\xx \cdot \dot{\nn})\, ds 
%    = \int_\gamma \left(\xx \cdot \dot{\xx}_s^\perp \right)
%    = -\int_\gamma \left(\xx^\perp \cdot \dot{\xx}_s\right)\, ds 
%    = \int_\gamma \left(\xx_s^\perp \cdot \dot{\xx}\right)\, ds 
%    = \int_\gamma (\dot{\xx} \cdot \nn)\, ds.
%\end{align}
\begin{align}
  \dot{A}(t) = \int_\gamma \left(\beta (\ff \cdot \nn)\nn 
    + \SS[\ff]\right) \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where we have used the incompressibility of the single-layer potential.
Repeatedly applying integration by parts, we have
\begin{align}
  \dot{A}(t) &= \beta \int_\gamma \left( -\xx_{ssss} + 
    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
  &= -\beta \int_\gamma \left(-\xx_{sss} + \sigma \xx_s 
    \right) \cdot \kappa \xx_s \, ds \\
  &= -\beta \int_\gamma \left(\xx_{ss} \cdot 
    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
  &= -\beta \int_\gamma \left(\kappa (\xx_{ss} \cdot \xx_{ss}) + 
    \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
    \, ds \\
  &= -\beta \int_\gamma \left(\kappa^3 + \kappa \sigma \right) 
    \, ds = \beta \int_\gamma \left(
    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
  \label{eqn:areaROC}
\end{align}
Note that the integrand in~\eqref{eqn:areaROC} is zero when the vesicle
reaches its steady state which confirms that $\dot{A}(t) = 0$.

>>>>

% In this work we use numerical simulations to examine the hydrodynamics
% of a semipermeable vesicle in two dimensions.

Vesicles (self-enclosed lipid bilayer membranes) have been
used as a model system to study reshaping, remodeling, and scission of
cell membranes due to osmotic stress~\cite{OgleckaEtAl2014_eLife,
CamposSaric2020_bioRxiv, Dimova2020_GVB, BhatiaChrist2020_SoftMatt}.
Under an osmotic shock, a vesicle may undergo oscillatory variation in
its enclosed volume~\cite{ChabanonHoLiedberg2017_BJ}. When trapped in a
microfluidic channel, an osmotic shock induces vesicle shape changes,
tank-treading and tumbling
motions~\cite{BhatiaRobinsonDimova2020_SoftMatt,
BhatiaChrist2020_SoftMatt}, an illustration that the permeable water
flow is inevitably coupled with the mechanical pressure jump across the
membrane~\cite{yao-mor2017}, which could become comparable to osmotic
stress when the membrane deformation is large~\cite{LiMoriSun2015_PRL,
yao-mor2017}.

In microfluidic experiments, it is challenging to keep a freely
suspended red blood cell or vesicle in a steady flow for more than a few
hours. It is also difficult to quantify the vesicle area/volume ratio
accurately.  Using numerical simulations we can quantify the effects of
water permeability on the hydrodynamics of a vesicle under mechanical
stresses. 
In the absence of an osmotic gradient, we first show that
a freely suspended semipermeable vesicle behaves differently from an
impermeable vesicle over a long time. Next we show that, at normal
physiological conditions, the permeable water flow can be amplified by
extreme confinement and results in observable change in both vesicle
hydrodynamics and reduced area. These results will shed light on the
dynamic aging process of living cells that constantly go through flows
and constrictions over a period of days.

@@

In experiments, vesicles under shear flows are measured for a minute
(chapter 19 in~\cite{Dimova2020_GVB}), a duration over which vesicle can
be assumed impermeable~\cite{AbkarianViallat2005_BJ}. Our simulations of
a semipermeable vesicle in these flows show that water permeability may
lead to different vesicle hydrodynamics over a time scale of 10 to 100
minutes. 
%{\bf YNY: is this correct? The shear flow results seem to suggest that
%the permeability may lead to change in area in less than a second???}
For vesicles in a narrow microfluidic channel, the experiments usually do not last longer than a few minutes~\cite{abk-fai-sto2006}.
%\todo[inline]{Howard talked aboutthis statement not always being true}. 
Our simulations show that strong
confinement can lead to amplification of water permeability at such
short time scale.
%
%all vesicles have a length
%$L$, but the area is not conserved because of semipermeability. We vary
%the initial vesicle reduced area, the permeability rate, and the flow
%rate. For all simulations, the bending modulus is $10^{-19}J$.

@@

Semipermeability is introduced by modifying the vesicle velocity from
the standard no-slip condition $\dot{\xx} = \uu(\xx)$. Instead,
the permeating water flux through the lipid bilayer membrane is proportional to the
membrane traction~\cite{yao-mor2017}. That is, 
\begin{align}
  \label{eqn:vesicleFlux}
  \uu - \dot{\xx} = - k_w (\ff_\mathrm{mem} \cdot \nn) \nn, \qquad
  \xx \in \gamma,
\end{align}
where $k_w$ is the water permeability.

@@

We next compute the transient area of the vesicle. Since the area of the
vesicle is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds,
  \label{eqn:area}
\end{align}
its derivative includes terms due to the time derivative of $\xx \cdot
\nn$ and a term due to interface stretching~\cite{lai-tse-hua2008}, but
this latter term term is zero because of the local inextensibility
condition. Therefore,
\begin{align}
  \dot{A}(t) =
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
Applying integration by parts, the two integrals agree, and we have
\begin{align}
  \dot{A}(t) = \int_\gamma \left(\beta (\ff \cdot \nn)\nn 
    + \SS[\ff]\right) \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where we have used the incompressibility of the single-layer potential.
Repeatedly applying integration by parts, we have
\begin{align}
  \dot{A}(t) &= \beta \int_\gamma \left( -\xx_{ssss} + 
    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
  &= -\beta \int_\gamma \left(-\xx_{sss} + \sigma \xx_s 
    \right) \cdot \kappa \xx_s \, ds \\
  &= -\beta \int_\gamma \left(\xx_{ss} \cdot 
    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
  &= -\beta \int_\gamma \left(\kappa (\xx_{ss} \cdot \xx_{ss}) + 
    \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
    \, ds \\
  &= -\beta \int_\gamma \left(\kappa^3 + \kappa \sigma \right) 
    \, ds = \beta \int_\gamma \left(
    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
  \label{eqn:areaROC}
\end{align}
Note that the integrand in~\eqref{eqn:areaROC} is zero when the vesicle
reaches its steady state which confirms that $\dot{A}(t) = 0$.

\section*{Vesicle Tensions}
\begin{figure*}[htp]
  \centering
  \ifTikz
    \input{figures/relaxationTensions.tikz}
  \else
    \includegraphics{figures/relaxationTensions.pdf}
  \fi
  \caption{\label{fig:relaxationTensions} Tension of images in
  relaxation.}
\end{figure*}
